\center \fontsize{18.8}{20} \selectfont Towards Globally Color-Dual All-Loop Integrands

### 2023-08-31 notes/2loopNLSM.tex
Finished first pass on unitarity rework.

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes, it is
convenient to project onto a specific basis of the Del
Duca-Dixon-Maltoni (DDM) color basis \cite{DixonMaltoni} by only
inserting color-ordered amplitudes in the product of trees, and
restricting the set of compatible diagrams to those whose color
factors reduce to the appropriate DDM element when summing over
\emph{only the uncut color contractions}.  In this situation, the
color becomes an overall factor on the entire equality and thus can be
ignored.  Cuts that are organized in this manner are known as
``color-ordered'' cuts, in the same sense as color-ordered tree
amplitudes.  Since the DDM basis (and dual Kleiss-Kuijf amplitude
basis \cite{Kleiss:1988ne}) have $(n-2)!$ elements, one generally
needs to evaluate that many separate color-orderings of the same cut.
As an example, consider the ``kissing triangle'' cut topology, one of
the factorization channels of the iterated unitarity cut, which has
two topologically distinct color-ordered expansions
\begin{equation}
  \label{eq:kt-color-ord}
  \KissingTriangles =
  \begin{cases}
     s\text{-}t \text{ channel}
    &
      \frac{1}{\textcolor{ucp-color}{s_{12}}}
      n \left[\scalebox{0.65}{\KTBTContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]\\
    t\text{-}u \text{ channel}
    & 
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
      n\left[
      \scalebox{0.65}{\KTDBContrib{1}{3}{4}{2}}
      \right]
  \end{cases}
\end{equation}
which serve as a complete basis for the full cut.

Even beginning at one loop, we will encounter types of cuts that are
difficult to properly define due to hidden non-physical singularities.
The quintessential example of such a ``pathological'' cut is the 3-1
two-particle cut at one loop which hides a bubble-on-external-leg
singularity
\begin{equation}
  \begin{tikzpicture}
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {};
\vertex (a2) at (-.8, 0.6) {};
\vertex (a3) at (-1, 0) {};
\vertex [dot, scale=2](mid1) at (0,0){};
\vertex [dot, scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot, scale=2](mid3) at (1,0){};
\vertex [dot, scale=1.5,hgrey0](mid4) at (1,0){};
\vertex (a4) at (1.8, 0) {};
\diagram{
(mid1) --[ultra thick,](a1),
(mid1) --[ultra thick,](a2),
(mid1) --[ultra thick,](a3),
%(mid1) --[ultra thick,](mid3),
(mid1) --[ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[ultra thick](a4),
};
\end{feynman}
\end{tikzpicture}
\ni
\frac{\delta(\textcolor{cut-color}{k_1^2})}{\textcolor{ucp-color}{k_1^2}}
n \left[
  \begin{tikzpicture}
    \ver{e1}{-1,0};
    \ver{e2}{-0.6,0.4};
    \ver{e3}{-0.3,0.4};
    \ver{i1}{-0.6,0};
    \ver{i2}{-0.3,0};
    \ver{i3}{0,0};
    \ver{i4}{0.8,0};
    \ver{co}{$(i3)!0.5!(i4)$};
    \ver{e4}{1.2,0};
    \draw[ultra thick] (e1)--(i1)--(i2)--(i3) to [out=80,in=100,min distance=0.4cm](i4)--(e4);
    \draw[ultra thick,ucp-color](i1)--(i2)--(i3);
    \draw[ultra thick,cut-color](i4)--(e4);
    \draw[ultra thick] (i3) to [out=-80,in=-100,min distance=0.4cm](i4);
    \draw[ultra thick] (e2) -- (i1);
    \draw[ultra thick] (e3)--(i2);
    \draw[ultra thick,dashed,cut-color] (co)--++(90:0.5);
    \draw[ultra thick,dashed,cut-color] (co)--++(-90:0.5);
  \end{tikzpicture}
  \right]
\end{equation}

The essence of generalized unitarity is that when internal lines are
placed on-shell, the integrand should factorize into a product of
on-shell tree amplitudes.  The cut to be performed can, itself, be
encoded in a graph.  For example, potential 4pt 2-loop cuts include
\begin{equation}
\label{eq:PotentialCuts}
\text{(a)} \SubtleCut ~~~~ \text{(b)} \PhysicalCutOne{}{}{}{} ~~~~ \text{(c)} \PhysicalCutTwo{}{}{}{} ~~~~ \text{(d)}  \MaxCut
\end{equation}
as well as many more.
In these graphs, all exposed propagators are understood to be cut/on-shell.
As an equation, a unitarity cut looks schematically like
\begin{equation}
\label{eq:SchematicCutEquation}
\sum \limits_\text{topologies} \frac{n \vert_\text{cut}}{\prod \limits_\text{uncut} k^2} = \sum \limits_\text{states} \prod \limits_\text{blobs} A_\text{tree} \, .
\end{equation}
The right hand side of the equation is a product of on-shell tree
amplitudes, represented graphically in \cref{eq:PotentialCuts} by grey
blobs.  States crossing (or propagating across) the cut must be summed
over.  For scalars this sum is trivial but for spinning particles this
is no longer the case and the physical state projector must be
inserted (see Dimitrius paper for simplifying this operation).  The
graph topologies appearing in the left hand side of
\cref{eq:SchematicCutEquation} come from blowing up the grey blobs
into cubic tree graphs in every conceivable way.  The numerator of
each graph is related back to the basis numerators via Jacobi
relations.  For each of these topologies, the numerator of the
associated graph is subjected to the cut constraints and is then
divided by the uncut propagators.  The right hand side is assumed to
be known, which places a constraint on the numerator ansatz in the
left hand side.

@@

In our construction, only well-defined cuts are imposed so, for
example, cut (a) of \cref{eq:PotentialCuts} is not enforced since it
is possible to blow up the left blob into a tree amplitude with an
on-shell internal propagator.  Cuts (b), (c), and (d) represent
physical cuts where (d) is known as a maximal cut since the maximal
number of propagators have been put on shell.

@@

Proceeding to the next-to-maximal cuts, we continue to avoid BEL
diagrams.  Additionally, now that we have higher-point vertices, we
avoid diagrams in which the higher-point vertex could contain an
off-shell propagator that carries the same momenta as one of the cut
propagators, \eg \ace{Any feedback on how this looks, especially with
  the dashed lines?} \jm{It looks good!  Can you omit the part where the blue lines actually cross?}\ace{Like so?}
\begin{equation*}
  \begin{tikzpicture}
    \pgfmathsetmacro{\ri}{1};
    \pgfmathsetmacro{\re}{0.6};
    \pgfmathsetmacro{\ang}{35};
    \begin{feynman}
      \blob{i1}{\ang:\ri};
      \blob{i2}{180-\ang:\ri};
      \blob{i3}{180+\ang:\ri};
      \blob{i4}{-\ang:\ri};
      \blob{i5}{ $(i1)!0.5!(i2)$ };
      \ver{e1}{45:\ri+\re};
      \ver{e2}{135:\ri+\re};
      \ver{e3}{225:\ri+\re};
      \ver{e4}{-45:\ri+\re};

      \ver{c1}{0:\cl};
      \ver{c2}{\cang:\cl};
      \ver{c3}{2*\cang:\cl};
      \ver{c4}{3*\cang:\cl};
      \ver{ci}{$(i5)!0.5!(i1)$};
      \path let \p1 =(ci) in coordinate (c5) at (\x1,0.2){};
      \draw[ultra thick] (e1)--(i1);
      \draw[ultra thick] (e2)--(i2);
      \draw[ultra thick] (e3)--(i3);
      \draw[ultra thick] (e4)--(i4);
      \draw[ultra thick] (i5)--(i1)--(i4)--(i3)--(i2)--(i6);
      \draw[ultra thick] (i6) to[out=60,in=180-60] (i5);
      \draw[ultra thick] (i6) to[out=-60,in=180+60] (i5);
      \draw[dashed,ultra thick,cut-color] (0,0) -- (c1);
      \draw[dashed,ultra thick,cut-color] (0,0) -- (c2);
      \draw[dashed,ultra thick,cut-color] (0,0) -- (c3);
      \draw[dashed,ultra thick,cut-color] (0,0) -- (c4);
      \draw[dashed,ultra thick,cut-color] (c5) -- ++(90:0.7*\cl);
      \node[fill=white,circle] (c) at (0,0) {};

  \end{tikzpicture},
\end{equation*}
as it is non-trivial and beyond the scope of this paper to define
exactly what such a cut and cut contribution means.  After discarding
these types of cuts, we are left with 8 unique cut topologies, shown
in \cref{fig:ym-nmc}, to check.  Seven of these next-to-maximal cuts
are consistent with the kinematic Jacobi relations and symmetries.
Interestingly, the ``kissing triangles'' cut,
\begin{equation*}
  \KissingTriangles \,,
\end{equation*}
cannot be satisfied by either the unrestricted or the BDN-restricted
ansatz with Jacobi relations and symmetries applied.

>>>>

In color-charged theories, both sides of the cut can be further
decomposed according to the color algebra.  For theories charged under
the adjoint of $SU(N)$ via $f^{abc}$-dressed amplitudes, it is
convenient to project onto a specific element of the Del
Duca-Dixon-Maltoni (DDM) color basis \cite{DixonMaltoni} by only
inserting color-ordered amplitudes in the product of trees, and
restricting the set of compatible diagrams to those whose color
factors reduce to the appropriate DDM element when summing over
\emph{only the uncut color contractions}.  In this situation, the
color becomes an overall factor on the entire equality and thus can be
ignored.  Cuts that are organized in this manner are known as
``color-ordered'' cuts, in the same sense as color-ordered tree
amplitudes.  Since the DDM basis (and dual Kleiss-Kuijf amplitude
basis \cite{Kleiss:1988ne}) have $(n-2)!$ elements, one generally
needs to evaluate that many separate color-orderings of the same cut.
As an example, consider the ``kissing triangle'' cut topology, one of
the factorization channels of the iterated unitarity cut, which has
two topologically distinct color-ordered expansions
\begin{equation}
  \label{eq:kt-color-ord}
  \KissingTriangles =
  \begin{cases}
     s\text{-}t \text{ channel}
    &
      \frac{1}{\textcolor{ucp-color}{s_{12}}}
      n \left[\scalebox{0.65}{\KTBTContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]\\
    t\text{-}u \text{ channel}
    & 
      \frac{1}{\textcolor{ucp-color}{\ell_1^2}} \
      n \left[
      \scalebox{0.65}{\KTDBContrib{1}{4}{3}{2}}
      \right]
      +
      \frac{1}{\textcolor{ucp-color}{\ell_2^2}} \
      n\left[
      \scalebox{0.65}{\KTDBContrib{1}{3}{4}{2}}
      \right]
  \end{cases}
\end{equation}
which serve as a complete basis for the full cut.

Even beginning at one loop, we encounter classes of cuts that are
difficult to properly define due to hidden non-physical singularities.
The quintessential example of such a ``pathological'' cut is the 3-1
two-particle cut at one loop which hides a bubble-on-external-leg
singularity
\begin{equation}
  \begin{tikzpicture}
\begin{feynman}
\vertex (a1) at (-.8, -0.6) {1};
\vertex (a2) at (-.8, 0.6) {3};
\vertex (a3) at (-1, 0) {2};
\vertex [dot, scale=2](mid1) at (0,0){};
\vertex [dot, scale=1.5,hgrey0](mid2) at (0,0){};
\vertex [dot, scale=2](mid3) at (1,0){};
\vertex [dot, scale=1.5,hgrey0](mid4) at (1,0){};
\vertex (a4) at (1.8, 0) {4};
\diagram{
(mid1) --[ultra thick,](a1),
(mid1) --[ultra thick,](a2),
(mid1) --[ultra thick,](a3),
%(mid1) --[ultra thick,](mid3),
(mid1) --[ultra thick,out=60,in=120,min distance=0.4cm](mid3),
(mid1) --[ultra thick,out=-60,in=-120,min distance=0.4cm](mid3),
(mid3) --[ultra thick](a4),
};
\end{feynman}
\end{tikzpicture}
\ni
\frac{\delta(\textcolor{cut-color}{k_4^2})}{\textcolor{ucp-color}{s\, k_4^2}}
n \left[
  \begin{tikzpicture}
    \ver{e1}{-1,0};
    \ver{e2}{-0.6,0.4};
    \ver{e3}{-0.3,0.4};
    \ver{i1}{-0.6,0};
    \ver{i2}{-0.3,0};
    \ver{i3}{0,0};
    \ver{i4}{0.8,0};
    \ver{co}{$(i3)!0.5!(i4)$};
    \ver{e4}{1.2,0};
    \draw[ultra thick] (e1)--(i1)--(i2)--(i3) to [out=80,in=100,min distance=0.4cm](i4)--(e4);
    \draw[ultra thick,ucp-color](i1)--(i2)--(i3);
    \draw[ultra thick,cut-color](i4)--(e4);
    \draw[ultra thick] (i3) to [out=-80,in=-100,min distance=0.4cm](i4);
    \draw[ultra thick] (e2) -- (i1);
    \draw[ultra thick] (e3)--(i2);
    \draw[ultra thick,dashed,cut-color] (co)--++(90:0.5);
    \draw[ultra thick,dashed,cut-color] (co)--++(-90:0.5);
  \end{tikzpicture}
  \right]
\end{equation}
in which one of the uncut propagators is set equal to an external
on-shell propagator.  Discussions about how to resolve these types of
singularities can be found in Refs.~\cite{Edison:2022jln,
  Edison:2022smn, Edison:2020uzf, Geyer:2015bja, Bern:2012uf}
\citepls{Minahan}.  In addition to the bubble-on-external-leg cuts,
there is a related class of cuts in which an uncut propagator is set
to zero by an \emph{internal cut condition}, for instance,
\begin{equation*}
  \begin{tikzpicture}
    \pgfmathsetmacro{\ri}{1};
    \pgfmathsetmacro{\re}{0.6};
    \pgfmathsetmacro{\ang}{35};
    \begin{feynman}
      \blob{i1}{\ang:\ri};
      \blob{i2}{180-\ang:\ri};
      \blob{i3}{180+\ang:\ri};
      \blob{i4}{-\ang:\ri};
      \blob{i5}{ $(i1)!0.5!(i2)$ };
      \ver{e1}{45:\ri+\re};
      \ver{e2}{135:\ri+\re};
      \ver{e3}{225:\ri+\re};
      \ver{e4}{-45:\ri+\re};

      \ver{c1}{0:\cl};
      \ver{c2}{\cang:\cl};
      \ver{c3}{2*\cang:\cl};
      \ver{c4}{3*\cang:\cl};
      \ver{ci}{$(i5)!0.5!(i1)$};
      \path let \p1 =(ci) in coordinate (c5) at (\x1,0.2){};
      \draw[ultra thick] (e1)--(i1);
      \draw[ultra thick] (e2)--(i2);
      \draw[ultra thick] (e3)--(i3);
      \draw[ultra thick] (e4)--(i4);
      \draw[ultra thick] (i5)--(i1)--(i4)--(i3)--(i2)--(i6);
      \draw[ultra thick] (i6) to[out=60,in=180-60] (i5);
      \draw[ultra thick] (i6) to[out=-60,in=180+60] (i5);
      \draw[dashed,ultra thick,cut-color] (0,0) -- (c1);
      \draw[dashed,ultra thick,cut-color] (0,0) -- (c2);
      \draw[dashed,ultra thick,cut-color] (0,0) -- (c3);
      \draw[dashed,ultra thick,cut-color] (0,0) -- (c4);
      \draw[dashed,ultra thick,cut-color] (c5) -- ++(90:0.7*\cl);
      \node[fill=white,circle] (c) at (0,0) {};
  \end{tikzpicture}.
\end{equation*}
Fully resolving both of these problems is beyond the scope of the
current work, so we will simply not impose any constraints that
require handling these type of cuts.

\iffalse
The essence of generalized unitarity is that when internal lines are
placed on-shell, the integrand should factorize into a product of
on-shell tree amplitudes.  The cut to be performed can, itself, be
encoded in a graph.  For example, potential 4pt 2-loop cuts include
\begin{equation}
\label{eq:PotentialCuts}
\text{(a)} \SubtleCut ~~~~ \text{(b)} \PhysicalCutOne{}{}{}{} ~~~~ \text{(c)} \PhysicalCutTwo{}{}{}{} ~~~~ \text{(d)}  \MaxCut
\end{equation}
as well as many more.
In these graphs, all exposed propagators are understood to be cut/on-shell.
As an equation, a unitarity cut looks schematically like
\begin{equation}
\label{eq:SchematicCutEquation}
\sum \limits_\text{topologies} \frac{n \vert_\text{cut}}{\prod \limits_\text{uncut} k^2} = \sum \limits_\text{states} \prod \limits_\text{blobs} A_\text{tree} \, .
\end{equation}
The right hand side of the equation is a product of on-shell tree
amplitudes, represented graphically in \cref{eq:PotentialCuts} by grey
blobs.  States crossing (or propagating across) the cut must be summed
over.  For scalars this sum is trivial but for spinning particles this
is no longer the case and the physical state projector must be
inserted (see Dimitrius paper for simplifying this operation).  The
graph topologies appearing in the left hand side of
\cref{eq:SchematicCutEquation} come from blowing up the grey blobs
into cubic tree graphs in every conceivable way.  The numerator of
each graph is related back to the basis numerators via Jacobi
relations.  For each of these topologies, the numerator of the
associated graph is subjected to the cut constraints and is then
divided by the uncut propagators.  The right hand side is assumed to
be known, which places a constraint on the numerator ansatz in the
left hand side.

@@

In our construction, only well-defined cuts are imposed so, for
example, cut (a) of \cref{eq:PotentialCuts} is not enforced since it
is possible to blow up the left blob into a tree amplitude with an
on-shell internal propagator.  Cuts (b), (c), and (d) represent
physical cuts where (d) is known as a maximal cut since the maximal
number of propagators have been put on shell.
\fi

@@

Proceeding to the next-to-maximal cuts, we continue to avoid BEL
diagrams.  Additionally, now that we have higher-point vertices, we
avoid diagrams in which the higher-point vertex could contain an
off-shell propagator that carries the same momenta as one of the cut
propagators, \eg \ace{Any feedback on how this looks, especially with
  the dashed lines?} \jm{It looks good!  Can you omit the part where the blue lines actually cross?}\ace{Like so?}

as it is non-trivial and beyond the scope of this paper to define
exactly what such a cut and cut contribution means.  After discarding
these types of cuts, we are left with 8 unique cut topologies, shown
in \cref{fig:ym-nmc}, to check.  Seven of these next-to-maximal cuts
are consistent with the kinematic Jacobi relations and symmetries.
Interestingly, the ``kissing triangles'' cut,
\begin{equation*}
  \KissingTriangles \,,
\end{equation*}
cannot be satisfied by either the unrestricted or the BDN-restricted
ansatz with Jacobi relations and symmetries applied.
